---
title: "Effect of Population Density and Land Use on Fukushima Air Dose Rate Reduction"
author:
- Student; Jovani Ntabgoba
- Supervisor; Prof Atsushi Togashi
- Assistant Supervisor; Prof Suguri Hiroki
date: "August 10, 2016"
output: html_document
---

## DOES POPULATION DENSITY AND LAND USE AFFECT THE RATE OF REDUCTION OF FUKUSHIMA DAICHI RADIATIONS? 



## Part I: Data Loading, Cleaning and Exploring

#### 1.1 Nihomatsu City,70km to Fukushima Daichi

1.11 Required Packages
```{r, error=FALSE,warning=FALSE, message=FALSE}
library("leaflet")
library(readr)
library(dplyr)
library(RColorBrewer)
library(Hmisc)
library(ggplot2)
library(formatR)
```


##### 1.11 Loading June 2011 and Nov 2013 Fukushima Data and selecting Nihomatsu's.
```{r}
fuk2011 <- read.csv("jun_2011_fukushima.csv")
fuk2013 <- read.csv("nov_2013_fukushima.csv")
```

##### 1.12 Change to machine readeable column names
```{r, warning = FALSE, message = FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
names(fuk2011) <- c("gridcode","pref","city","gridCenterNorthlat","gridCenterEastlng","gridCenterNorthlatDec",
                 "gridCenterEastlngDec","daichi_distance","no_samples","AvgAirDoseRate",
                       "NE_nLat","NE_eLong","NW_nLat","NW_eLong",
                       "SW_nLat","SW_eLong","SE_nLat","SE_eLong")
names(fuk2013) <- c("gridcode","pref","city","gridCenterNorthlat","gridCenterEastlng","gridCenterNorthlatDec",
                     "gridCenterEastlngDec","daichi_distance","no_samples","AvgAirDoseRate",
                     "NE_nLat","NE_eLong","NW_nLat","NW_eLong",
                     "SW_nLat","SW_eLong","SE_nLat","SE_eLong")
dim(fuk2013); dim(fuk2011)
max(fuk2013$AvgAirDoseRate);min(fuk2013$AvgAirDoseRate)
max(fuk2011$AvgAirDoseRate);min(fuk2011$AvgAirDoseRate)
```


##### 1.14 Create air dose quantiles that are plot-able,i.e 6 categorical variables.
```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
fuk2011_q <- fuk2011 %>%
        mutate(dose_quants = cut2(fuk2011$AvgAirDoseRate,
                                  cuts=c(0.1,1.0,5.0,15.0,25.0,35.0,45.0),
                                  levels.mean=TRUE))
fuk2013_q <- fuk2013 %>%
        mutate(dose_quants = cut2(fuk2013$AvgAirDoseRate,cuts=c(0.1,1.0,5.0,15.0,25.0,35.0,45.0),levels.mean=TRUE))
```

<ul>
 <li>Visible reduction of Average Air Dose Distribution by half  in Nihomatsu.</li>
 <li>Trouble is knowing the distribution of causes of this reduction?</li>
</ul>

##### 1.15 Color function
```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
iro <- colorFactor(
        palette = "YlOrRd",
        domain = fuk$dose_quants
)
iro2013 <- colorFactor(
        palette = "YlOrRd",
        domain = fuk2013_q$dose_quants
)
# Link of Daichi
fukulink <- paste(sep = "<br/>",
                  "<br><a href='http://www.tepco.co.jp
                  /en/decommision/index-e.html'>Fukushima Daichi</a></b>",
                  "Source of radiations")
```

##### 1.16 Nihomatsu Average Air Dose Rate for 2011
```{r, warning = FALSE, message = FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
niho_plot <- leaflet() %>%
        addTiles()%>%
        addRectangles(data = niho_q,lng1 = ~SW_eLong, lat1 = ~SW_nLat,
                      lng2 = ~NE_eLong, lat2 = ~NE_nLat,
                      color = ~iro(niho_q$dose_quants))%>%
        addLegend("bottomright", pal = iro, values = niho_q$dose_quants,
                  title = "AvgAirDoseRate",
                  labFormat = labelFormat(prefix = "µSv/h "),
                  opacity = 1)%>%
        addPopups(lat = 37.4211, lng = 141.0328,popup = fukulink,
                  options = popupOptions(closeButton = TRUE)) 
niho_plot
```
<br>

##### 1.17 Nihomatsu Average Air Dose Rate for 2013
```{r, warning = FALSE, message = FALSE}
niho2013_plot <- leaflet() %>%
        addTiles()%>%
        addRectangles(data = nihom2013_q,lng1 = ~SW_eLong, lat1 = ~SW_nLat,
                      lng2 = ~NE_eLong, lat2 = ~NE_nLat,
                      color = ~iro2013(nihom2013_q$dose_quants))%>%
        addLegend("bottomright", pal = iro2013, values = nihom2013_q$dose_quants,
                  title = "AvgAirDoseRate",
                  labFormat = labelFormat(prefix = "µSv/h "),
                  opacity = 1)%>%
        addPopups(lat = 37.4211, lng = 141.0328,popup = fukulink,
                  options = popupOptions(closeButton = TRUE)) 
niho2013_plot
```
<br>

##### 1.18 Nihommatsu 2011, Counts of Measuring Locations per Air Dose Rate
```{r, warning = FALSE, message = FALSE}
ggplot(niho_q, aes(daichi_distance,AvgAirDoseRate)) +
        geom_point() +
        geom_smooth(se = FALSE)+
        ggtitle("AvgAirDose against Distance to Daichi Plant")
```

##### 1.19 Nihommatsu 2011,Counts of Measuring Locations per Air Dose Rate 
```{r, warning = FALSE, message = FALSE}
ggplot(data = niho_q) +
        geom_bar(mapping = aes(x = daichi_distance, fill = dose_quants), width = 1)+
        ggtitle("AvgAirDose Measured Counts against Daichi Distance")
```

#### 1.2 Minamisoma City, 30km to Fukushima Daichi
```{r, warning = FALSE, message = FALSE}
mina <- read_csv("niho.csv")
dim(mina)
```

##### 1.21 Change to machine readeable column names
```{r, warning = FALSE, message = FALSE}
#change Minamisoma city to minamisoma
mina$city[mina$city == "Minamisoma city"] <- "minamisoma"
#filter nihomatsu observations only
mina <- subset(mina, city == "minamisoma")
dim(mina)
summary(mina$AvgAirDoseRate)
```

##### 1.22 Create air dose quantiles that are plot-able,6 categorical variables.
```{r, warning = FALSE, message = FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mina_q <- mina %>%
        mutate(mina_quants = cut2(mina$AvgAirDoseRate,cuts=seq(0.2,20,2.2),levels.mean=TRUE))
mina_q <- na.omit(mina_q) #remove NAs
write_csv(mina_q, path = "mina_q.csv")
```

##### 1.23 Color function
```{r, warning = FALSE, message = FALSE}
mina_iro <- colorFactor(
        palette = "Set1",
        domain = mina_q$mina_quants
)
```

##### 1.24 Minamisoma Average Air Dose Rate for 2011
```{r,warning = FALSE, message = FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mina_plot <- leaflet() %>%
        addTiles()%>%
        addRectangles(data = mina_q,lng1 = ~SW_eLong, lat1 = ~SW_nLat,
                      lng2 = ~NE_eLong, lat2 = ~NE_nLat,
                      color = ~mina_iro(mina_q$mina_quants))%>%
        addLegend("bottomright", pal = mina_iro, values = mina_q$mina_quants,
                  title = "AvgAirDoseRate",
                  labFormat = labelFormat(prefix = "µSv/h "),
                  opacity = 1)%>%
        addPopups(lat = 37.4211, lng = 141.0328,popup = fukulink,
                  options = popupOptions(closeButton = TRUE)) 
mina_plot
```

<br>

##### 1.25 Minamisoma AvgAirDose against Distance to Daichi Plant
```{r, warning = FALSE, message = FALSE}
ggplot(mina_q, aes(daichi_distance,AvgAirDoseRate)) +
        geom_point() +
        geom_smooth(se = FALSE)+
        ggtitle("AvgAirDose against Distance to Daichi Plant")
```

##### 1.26 Minamisoma 2011, Counts of Measuring Locations per Air Dose Rate
```{r, warning = FALSE, message = FALSE}
ggplot(data = mina_q) +
        geom_bar(mapping = aes(x = daichi_distance, fill = mina_quants), width = 1)+
        ggtitle("Counts of Measuring Locations per Air Dose Rate")
```

## PART II: Calculating the Coefficient of Air Dose Reduction

#### 2.00 Merging 2011 and 2013 Nihomatsu Data Sets
```{r, warning = FALSE, message = FALSE,tidy=TRUE,fig.width=6, fig.height=4}
library(ggplot2)
niho_q <- read_csv("niho_q.csv")
niho2013_q <- read_csv("niho2013.csv")
names(niho_q) <- c("gridcode","pref","city","gridCenterNorthlat","gridCenterEastlng",
                   "gridCenterNorthlatDec","gridCenterEastlngDec","daichi_distance",
                   "no_samples","AvgAirDoseRate2011","NE_nLat","NE_eLong","NW_nLat","NW_eLong",
                   "SW_nLat","SW_eLong","SE_nLat","SE_eLong")
names(niho2013_q) <- c("gridcode","pref","city","gridCenterNorthlat","gridCenterEastlng",
                       "gridCenterNorthlatDec","gridCenterEastlngDec","daichi_distance",
                       "no_samples","AvgAirDoseRate2013","NE_nLat","NE_eLong","NW_nLat","NW_eLong",
                     "SW_nLat","SW_eLong","SE_nLat","SE_eLong")
niho11_13 <- merge(niho_q, niho2013_q, by.x = "gridcode", by.y = "gridcode", all = TRUE)
# Check if the merged columns are real identical
niho11_13 <- na.omit(niho11_13)
identical(niho11_13$gridCenterNorthlat.x,niho11_13$gridCenterNorthlat.y)
identical(niho11_13$daichi_distance.x,niho11_13$daichi_distance.y)
identical(niho11_13$no_samples.x,niho11_13$no_samples.y) #2011 and 2013 samples differ

niho11_13 <- select(niho11_13,gridcode,pref.x,city.x,gridCenterNorthlat.x,
                    gridCenterEastlng.x,gridCenterNorthlatDec.x,
                 gridCenterEastlngDec.x,daichi_distance.x,no_samples.x,AvgAirDoseRate2011,
                       NE_nLat.x,NE_eLong.x,NW_nLat.x,NW_eLong.x,no_samples.y,AvgAirDoseRate2013)
# create new data set (niho11_13)
write_csv(niho11_13,path = "niho11_13.csv")
niho11_13 <- read_csv("niho11_13.csv")
#Compare AvgAirDoseRate of 2011&2013
plot(x=niho11_13$AvgAirDoseRate2011,type="l",col="red",ylab = "Avg Air Dose Rate",
     xlab = "Frequency of measuring",main = "Compare AvgAirDoseRate of 2011&2013 Nihomatsu")
lines(niho11_13$AvgAirDoseRate2013,col="green")
legend("topright", legend = c("AvgAirDoseRate2011","AvgAirDoseRate2013"))
```

#### 2.10 Air Dose Rate Reduction due to Distance to Fukushima Daichi 

$$AvgAirDoseRate = \beta_{0} + \beta_{1} daichi distance + \epsilon_{i}$$

```{r, warning = FALSE, message = FALSE}
fit1 <- lm(AvgAirDoseRate2013~daichi_distance.x, data = niho11_13)
summary(fit1)
#Confidence interval
confint(fit1)
```


#### 2.20 Air Dose Rate Reduction due to Population Density and Land Use

$$AvgAirDoseRate = \beta_{0} + \beta_{1} popn density + \beta_{2} human activities + \epsilon_{i}$$

**Challenge**
<p> The Air Dose Rate measurements were collected from three sources: </p>
<ol>
 <li>Cars and Buses with Sensors: Measure Air of 100m&#0178 from the road</li>
 <li>Un manned Auto Helicopter: Measure from 300m above ground</li>
 <li>Fixed Sensors: Geographically sparsed </li>
</ol>

<p> What constitutes Land Use?: </p>
<ol>
 <li>Farming: Short life span crops lead to frequent land tilling</li>
 <li>Cleaned Places: Parks, GolfCourses, Gardens are cleaned too and could reduce radiations</li>
<br>
</ol>
<p> Population density and land usage is measured on a 500m&#0178; basis</p>
```{r, warning = FALSE, message = FALSE}
fuk_pop <- read.csv("fuk.csv")
View(fuk_pop)
# Convert mesh gridecode to lat/long
mesh_latlong <- function (mesh, loc = "C")
{
    mesh <- as.character(mesh)

    if (length(grep("^[0-9]{4}", mesh)) == 1) { 
        mesh12 <- as.numeric(substring(mesh, 1, 2))
        mesh34 <- as.numeric(substring(mesh, 3, 4))
        lat_width  <- 2 / 3;
        long_width <- 1;
    }
    else {
        return(NULL)
    }

    if (length(grep("^[0-9]{6}", mesh)) == 1) { 
        mesh5 <- as.numeric(substring(mesh, 5, 5))
        mesh6 <- as.numeric(substring(mesh, 6, 6))
        lat_width  <- lat_width / 8;
        long_width <- long_width / 8;
    }

    if (length(grep("^[0-9]{8}", mesh)) == 1) { 
        mesh7 <- as.numeric(substring(mesh, 7, 7))
        mesh8 <- as.numeric(substring(mesh, 8, 8))
        lat_width  <- lat_width / 10;
        long_width <- long_width / 10;
    }


    lat  <- mesh12 * 2 / 3;          
    long <- mesh34 + 100;

    if (exists("mesh5") && exists("mesh6")) {  
        lat  <- lat  + (mesh5 * 2 / 3) / 8;
        long <- long +  mesh6 / 8;
    }
    if (exists("mesh7") && exists("mesh8")) {  
        lat  <- lat  + (mesh7 * 2 / 3) / 8 / 10;
        long <- long +  mesh8 / 8 / 10;
    }

    if (loc == "C") {   
        lat  <-  lat  + lat_width  / 2;
        long <-  long + long_width / 2;
    }
    if (length(grep("N", loc)) == 1) {  
        lat  <- lat  + lat_width;
    }
    if (length(grep("E", loc) == 1)) {  
        long <- long +long_width;
    }

    lat  <- sprintf("%.8f", lat); 
    long <- sprintf("%.8f", long);

    x <- list(as.numeric(lat), as.numeric(long))
    names(x) <- c("lat", "long")

    return (x)
}
# j <- mesh_latlong(mesh = 564000463 , loc = "C")
# class(j);print(j);length(j)
grides <- fuk_pop[,1]
head(grides);class(grides);length(grides)
grides[2]
#create the lat/long
mylist <- list()
for (i in 1:length(grides) ){
        lis <- mesh_latlong(mesh = grides[i], loc = "C")
        mylist[[i]] <- lis
}

res <- as.data.frame(mylist)


#test
# df <- data.frame(matrix(unlist(res), nrow=10831, byrow=T))
# df <- ldply (res, data.frame)

library(data.table)
df <- as.data.frame(rbindlist(mylist, fill=TRUE))
df[,"gridcode"] <- grides 
View(df)
# merge gridcode and lat/lon datasets
fuk_ll <- merge(fuk_pop, df, by.x = "gridcode", by.y = "gridcode", all = TRUE)
write.csv(fuk_ll, file = "fuk_ll.csv")
#trial plots
pop_plot <- leaflet() %>%
        addTiles()%>%
        addPolylines(data = fuk_ll,lng = ~long, lat = ~lat)
pop_plot        
#remove duplicate lat/lon
uniq <- unique(x=df[,-3])

uniq_plot <- leaflet() %>%
        addTiles()%>%
        addCircles(data = uniq,lng = ~long, lat = ~lat)
uniq_plot

# Unique, lat or lon only
data <- subset(fuk_ll, !duplicated(fuk_ll[,7]))
dim(data)
View(data)
uniq_plot <- leaflet() %>%
        addTiles()%>%
        addCircles(data = data,lng = ~long, lat = ~lat)
uniq_plot
```

#### 2.30 Using other machine learning algorithms

<p> The Fukushima Nuclear Radiations is multi dimensional;</p>
<li> *Three major Isotopes each with;*</li>
<li> Time series due to half life </li>
<li> Magnitude of radiation (Becquerel) </li>
<li> Absorbed dose (Sievert (Sv)) </li>
<p> The above dimensionalities coupled with distance,population density and land use, create a data set that can be run on extensive machine learning algorithms like Support Vector Machines, Random Forest, Recurrent Neural Networks and more.</p>

**end**


